\section{Chapter 7}

%		7.15		%
\subsection*{7.15} Show that P is closed under the star operation. (Hint: Use dynamic programming. On input $y = y_1\cdots y_n$ for $y_i \in \Sigma$, build a table indicating for each $i \le j$ whether the substring $y_i\cdots y_j \in A^*$ for any $A \in$ P.)
\\
\textbf{Solution:} Let $M$ be a TM that decides $A$ in polynomial time. We construct a TM $T$ to decide $A^*$ in polynomial time:
\\ \\
$T$ = ``On input $w \in \Sigma^*$:
\begin{enumerate}
\item[1.]Let $|w| = n \in \mathbb{N}$.
\item[2.]If $n = 0$, $accept$ $w$.
\item[3.]Construct an $n \times n$ table $B$, initializing all elements to 0.
\item[4.]For $1 \le i \le n$ do:
\begin{enumerate}
\item[a.]If $w_i \in A$, set $B(i, i) = 1$.
\end{enumerate}
\item[5.]For $2 \le j \le n$ do:
\begin{enumerate}
\item[a.]For $1 \le k \le n-j+1$ do:
\begin{enumerate}
\item[i.]Set $m \leftarrow j+k-1$.
\item[ii.]If $w_k\cdots w_m \in A$, set $B(k,k) = 1$.
\item[iii.]For $k \le p \le m-1$ do:
\begin{enumerate}
\item[1.]If $B(k, p) = B(p, m) = 1$, set $B(k, m) = 1$.
\end{enumerate}
\end{enumerate}
\end{enumerate}
\item[6.]Output if $B(1, n) = 1$."
\end{enumerate}

%		7.20		%
\subsection*{7.20} We generally believe that \emph{PATH} is not NP-complete. Explain the reason behind this belief. Show that proving \emph{PATH} is not NP-complete would prove P $\ne$ NP.
\\
\textbf{Solution:} We can prove the converse: if P = NP, then \emph{PATH} is NP-complete. We will do two things (in assuming P = NP): show that \emph{PATH} is NP-hard, and because \emph{PATH} $\in$ NP, \emph{PATH} is NP-complete. We will do this by showing $SAT \le_p PATH$.

\par Since P = NP, there is a decider $S$ for $SAT$ that runs in time $O(n^k)$ for some $k \in \mathbb{N}$. We construct a TM $T$ that computes the reduction:
\\ \\
$T$ = ``On input \angles{\phi} where $\phi$ is a $SAT$ formula:
\begin{enumerate}
\item[1.]Run $S$ on input \angles{\phi}. If $S$ accepts \angles{\phi}, the construct a graph $G = (V, E)$ such that $s, t \in V, (s, t) \in E$.
\item[2.]If $S$ rejects \angles{\phi}, construct a graph $G = (V, E)$ such that $V = \{s, t\}, E = \emptyset$.
\item[3.]Output \angles{G, s, t}."
\end{enumerate}
Since this reduction takes polynomial time, then P = NP implies \emph{PATH} is NP-complete.

%		7.22		%
\subsection*{7.22} Let $DOUBLE$-$SAT = \{$\angles{\phi} $|$ $\phi$ has at least two satisfying assignments\}. Show that $DOUBLE$-$SAT$ is NP-complete.
\\
\textbf{Solution:} We will reduce $3SAT$ to this problem. Given $\phi$ as a $3SAT$ formula, all we need to do is to introduce a new variable $x$, and create a new formula $\phi'$ as follows:
\begin{center}
$\phi' = \phi \wedge (x \vee \overline{x})$
\end{center}
We show that $\phi \in 3SAT$ if and only if $\phi' \in DOUBLE$-$SAT$. If $\phi \in 3SAT$, then there is one satisfying assignment - for $\phi'$ there are two since we can set $x = 0, 1$ and still have $\phi'$ be true. If $\phi$ is unsatisfiable, then clearly $\phi'$ is also unsatisfiable. Therefore, we are done.