\section{Chapter 6}
\begin{enumerate}
%		6.6		%
\item[6.6]Describe two different Turing machines, $M$ and $N$, where $M$ outputs \angles{N} and $N$ outputs \angles{M}, when started on any input.
\\
\textbf{Solution:} Construct $M$ as follows:
\\
$M$ = ``On input $w$:
\begin{enumerate}
\itemsep0em
\item[1.]Obtain \angles{M} by the recursion theorem.
\item[2.]Compute $q($\angles{M}$)$, and write the result to the tape.
\item[3.]Halt."
\end{enumerate}
Now let $N = q($\angles{M}$)$. The computation in Step 2 writes \angles{q($\angles{M}$)} = \angles{N}. $N$ writes \angles{M} to the tape, so the construction of $M$ and $N$ is correct.

%		6.13		%
\item[6.13]For each $m > 1$ let $\mathbb{Z}_m = \{0,1,2,\cdots,m-1\}$, and let $F_m = (\mathbb{Z}_m, +, \times)$ be the model whose universe is $\mathbb{Z}_m$ and that has relations corresponding to the $+$ and $\times$ relations computed modulo $m$. Show that for each $m$, the theory Th($F_m$) is decidable.
\\
\textbf{Solution:} we are given a formula $\phi = Q_1x_1\cdots Q_kx_k \psi(x_1, \cdots, x_n)$ (assuming the $Q_i$ are quantifiers, and $\psi$ has no quantifiers). We iteratively define $I_\ell$ for $0 \le \ell \le n$:
\begin{itemize}
\item $I_n(x_1, \cdots, x_n) = \psi(x_1, \cdots, x_n)$,
\item $I_{\ell-1}(x_1, \cdots, x_{\ell-1}) = \bigvee_{i=0}^{m-1} I_{i}(x_1, \cdots, x_{\ell-1}, i)$ if $Q_\ell$ is $\exists$,
\item $I_{\ell-1}(x_1, \cdots, x_{\ell-1}) = \bigwedge_{i=0}^{m-1} I_{i}(x_1, \cdots, x_{\ell-1}, i)$ if $Q_\ell$ is $\forall$.
\end{itemize}
We then output the value of $I_0$. By an induction argument, we correctly calculate the truth value of $\phi$. Therefore, Th($F_m$) is decidable. 


%		6.23		%
\item[6.23]Show that the function $K(x)$ is not a computable function.
\\
\textbf{Solution:} Suppose that $K(x)$ were computable. We now construct a TM $Z$:
\\
$Z$ = ``On any input:
\begin{enumerate}
\itemsep0em
\item[1.]Obtain \angles{Z} by the recursion theorem.
\item[2.]Compute $d = K($\angles{Z}$)$.
\item[3.]Enumerate strings $x \in \Sigma^*$ in any order until $K(x) > d$.
\item[4.]Output $x$."
\end{enumerate}
This a contradiction since we are outputting a value of $K(x)$ that is greater than the own Turing Machine's value. Therefore, $K(x)$ is not computable.

%		6.24		%
\item[6.24]Show that the set of incompressible strings is undecidable.
\\
\textbf{Solution:} Since the set of incompressible strings is an infinite set, this reduces to 6.25.

%		6.25		%
\item[6.25]Show that the set of incompressible strings contains no infinite subset that is Turing-recognizable.
\\
\textbf{Solution:} Suppose that such a subset were Turing-recognizable, and let $E$ be the enumerator, since Turing-enumerable languages are exactly the Turing-recognizable ones. We now construct a TM $Y$:
\\
$Y$ = ``On any input:
\begin{enumerate}
\itemsep0em
\item[1.]Obtain \angles{Y} by the recursion theorem.
\item[2.]Run $E$ until it outputs a string $x$ with $|x| > |$\angles{Y}$|$.
\item[3.]Output $x$."
\end{enumerate}
This a contradiction - therefore, no infinite subset of the set of incompressible strings is Turing-recognizable.

%		6.27		%
\item[6.27]Let $S = \{$\angles{M} $|$ $M$ is a TM and $L(M) = \{$\angles{M}\}\}. Show that neither $S$ nor $\overline{S}$ is Turing-recognizable.
\\
\textbf{Solution:} We will show that $A_{TM} \le_m S$ and $A_{TM} \le_m \overline{S}$, implying that neither $S$ nor $\overline{S}$ is Turing-recognizable. For $A_{TM} \le_m S$, given \angles{M, w} $\in A_{TM}$, construct a TM $T$ that, on input $x$, rejects if $x \ne$ \angles{T} and simulates $M$ on $w$ otherwise. Therefore, $L(T) = \{$\angles{T}\} if $M$ accepts $w$, and $\emptyset$ otherwise.

\par For $A_{TM} \le_m \overline{S}$, change $T$'s construction to be ``accepts" instead of ``rejects". 
\end{enumerate}